---
title: "Shiny app - prep"
author: "Amelia Ritger"
date: "2/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
library(janitor)
library(sf)
library(tmap)
library(vegan)
library(gt)
```

## FOR FRAP
```{r}
reef <- read_csv("FRAP-Mohawk.csv")

#Tidy up the data
reef_tidy <- reef %>%
  pivot_longer(`Neighbor 1`:`Neighbor 4`) %>% 
  mutate(keep = ifelse(name==`Already counted?`, name, "keep")) %>% #create a column that we can subset all rows in a plot based on the presence of focal phylum in the plot at least once
  filter(keep == "keep")
```

## Organize the data 
```{r}
#Read in data
reef <- read_csv("MBONReef_Histogram.csv")

#Tidy up the data
reef_tidy <- reef %>%
  clean_names() %>% #standardize names
  #rename(latitude=lat_start, longitude=long_start) %>% #rename latitude and longitude
  group_by(location) %>% #group by location to get average lat/long values for each location
  mutate(latitude=head(lat_end,1), longitude=head(long_end,1)) %>% #get average lat/long values (because that's how that works...)
  ungroup() %>% #really important, we don't want to confuse R!
  pivot_longer("annelida_cirriformia_luxuriosa":"substrate_amphipod_tube_complex") %>%  #make long form
  separate(name, into="phylum", sep="_", remove=FALSE) %>% #Add column for phylum name
  mutate(vectorized_name=str_split(name, pattern="_")) %>% #In case this is useful...
  filter(!phylum=="no") %>% #remove values related to data collection issue
  filter(!phylum=="substrate") #remove substrate values

#Because it's faster to do it outside tidyverse
reef_tidy$binary <- ifelse(reef_tidy$value>0, 1, 0) #Add presence/absence column
reef_tidy$species <- gsub("^[^_]*_","",reef_tidy$name, perl=TRUE) #Add column for species name
reef_tidy$longitude <- ifelse(reef_tidy$longitude<0, reef_tidy$longitude, -reef_tidy$longitude) #because all longitude values in this region should be negative - looking at you, Rodes... 

#Do some more tidying
reef_tidy <- reef_tidy %>%
  mutate(species = str_replace(species, pattern="tube worm", "tubeworm")) %>%  #replace "tube worm" with "tubeworm" for later grouping
  separate(species, into="genus", sep="_", remove=FALSE) %>% #Add column for genus name
  mutate(species=str_replace_all(species, "_", " "),
         species=str_to_sentence(species)) %>%  #Make species names actually look like species names
  filter(!str_detect(species, pattern="dead")) #Remove instances where organism is dead
  #mutate(species_clean = str_remove(species, pattern="unknown"))
  #replace(species_clean, "hey there", str_detect(species_clean, "sponge"))

#Now time for some massive if, else statements to group organisms not identified to species
reef_tidy <- reef_tidy %>% 
  mutate(grouped_species = ifelse(str_detect(species, pattern = " worm"), "Other worms", ifelse(str_detect(species, pattern = "phoronid"), "Other worms", ifelse(str_detect(species, pattern = "tubeworm"), "Tubeworms", ifelse(str_detect(species, pattern = "algae"), "Other Algaes", ifelse(str_detect(species, pattern = "Filamentous"), "Other Algaes", ifelse(str_detect(species, pattern = "turf"), "Other Algaes", ifelse(str_detect(species, pattern = "blade"), "Other Algaes", ifelse(str_detect(species, pattern = "tunicate"), "Other Tunicates", ifelse(str_detect(species, pattern = "anemone"), "Other anemones", ifelse(str_detect(species, pattern = "bryozoan"), "Other bryozoans", ifelse(str_detect(species, pattern = "White fan"), "Other bryozoans", ifelse(str_detect(species, pattern = "sponge"), "Other Sponges", ifelse(str_detect(species, pattern = "Orange encrusting"), "Other Sponges", ifelse(str_detect(species, pattern = "Haliclona sp"), "Other Sponges", ifelse(str_detect(species, pattern = "zigzag"), "Other Hydroids", species)))))))))))))))) %>%
  mutate(grouped_genus = ifelse(str_detect(grouped_species, pattern = "Other"), grouped_species, ifelse(str_detect(grouped_species, pattern = "orange"), grouped_species, ifelse(str_detect(grouped_species, pattern = "White"), grouped_species, ifelse(str_detect(grouped_species, pattern = "encrusting"), grouped_species, ifelse(str_detect(grouped_species, pattern = "zigzag"), grouped_species, ifelse(str_detect(grouped_species, pattern = "solitary"), grouped_species, ifelse(str_detect(grouped_species, pattern = "sectioned"), grouped_species, genus)))))))) %>% #do the same for genus
  mutate(grouped_genus = str_to_title(grouped_genus)) %>%  #capitalize genus name
  mutate(phylum = str_to_title(phylum)) %>%  #capitalize phylum name
  mutate(kingdom = "Animalia")

genusnames <- unique(reef_tidy$genus)
```

## Create different datasets for different uses

#### Subset for a phylum
```{r, message=FALSE, warning=FALSE}
reef_phylum <- reef_tidy %>%
  filter(binary > "0") %>% #filter out species not present
  mutate(focal_genus="Corynactis") %>% #pick a focal phylum (BASED ON INPUT)
  mutate(to_match = ifelse(grouped_genus==focal_genus, filename, "FALSE")) %>% #create a column that we can subset all rows in a plot based on the presence of focal phylum in the plot at least once
  filter(filename %in% to_match) %>% #if focal phylum is present, keep all observations of that plot ("filename")
  #filter(!phylum=="annelida") %>% #remove the focal phylum, because we don't need to see co-occurrence of focal phylum with focal phylum (BASED ON INPUT)
  filter(location=="Mohawk") %>% 
  distinct(filename, genus, .keep_all=TRUE) #remove duplicate values within the same plot
```

#### Subset for a location
```{r, message=FALSE, warning=FALSE}
reef_location <- reef_tidy %>%
  filter(binary > "0") %>% #filter out species not present
  mutate(focal_location="Rodes") %>% #pick a focal location (BASED ON INPUT)
  mutate(to_match = ifelse(location==focal_location, filename, "FALSE")) %>% #create a column that we can subset all rows in a plot based on the presence of focal phylum in the plot at least once
  filter(filename %in% to_match) %>% #if focal phylum is present, keep all observations of that plot ("filename")
  #filter(!phylum=="annelida") %>% #remove the focal phylum, because we don't need to see co-occurrence of focal phylum with focal phylum (BASED ON INPUT)
  filter(phylum=="annelida") %>%  #select only the coocurring phyla you want to look at (BASED ON INPUT)
  distinct(filename, .keep_all=TRUE) #remove duplicate values within the same plot

#Find average abundance values for each location
reef_summary <- reef_tidy %>%
  filter(binary > "0") %>% #filter out species not present
  group_by(location,phylum) %>% #group by location, then lat/long
  summarize(mean_count = mean(value), #get the mean count
            median_count = median(value),
            sd_count = sd(value), #get the s.d. count
            iqr = IQR(value), #get the interquartile range for the count
            sample_size = n())

#Create separate dataframe of just latitude, longitude, and locations (use for later plotting species diversity/richness at each location)
reef_location_unique <- reef_tidy %>% 
  distinct(location, latitude, longitude) %>% 
  st_as_sf(coords=c("longitude", "latitude"), crs=4326) #Make sticky geometries for lat/long

```

#### Find species diversity/richness for each site
```{r, message=FALSE, warning=FALSE}
#Prep data
reef_vegan <- reef_tidy %>% #named so because of the vegan package!
  group_by(location,grouped_species) %>% #group by location, then lat/long
  summarize(mean_count = mean(value)) %>%  #get the mean count
  select(location, grouped_species, mean_count) %>% 
  ungroup()

#Calculate species diversity and richness for each site
reef_vegan_subset <- reef_vegan %>%
  pivot_wider(names_from=grouped_species, values_from=mean_count) %>% 
  select(`Abietinaria spp`:`Zonaria farlowii`)

diversity <- diversity(reef_vegan_subset)
richness <- specnumber(reef_vegan_subset)

# #Create a function to generate a continuous color palette
# Pal <- colorRampPalette(c('red','blue'))
# #Add a column of color values based on the diversity values
# reef_vegan %>% 
#   mutate(color = Pal(10)[as.numeric(cut(reef_vegan$diversity,breaks = 10))])
# #or, 
# reef_vegan$color_new <- colorRampPalette(c("blue", "red"))(length(unique(reef_vegan$diversity)))
# 
# #Add this information to the vegan dataframe
# reef_vegan <- reef_location_unique %>% 
#   add_column(diversity, richness, color=Pal(10)[as.numeric(cut(diversity(reef_vegan_subset),breaks = 10))])
# # Or
# reef_vegan <- reef_location_unique %>% 
#   add_column(diversity, richness) %>% 
#   mutate(color_brandnew = colorRampPalette(c("yellow", "red"))(length(richness)))

reef_vegan <- reef_location_unique %>% 
  add_column(diversity, richness)
```

## Let's try plotting something

#### The mean counts of each phylum (or species) occurring at each location:
```{r}
ggplot(reef_summary, aes(x=phylum,y=mean_count)) +
  geom_point(color="black",
             size=2) +
  geom_errorbar(aes(x=phylum,
                    ymin=mean_count - sd_count,
                    ymax=mean_count + sd_count), 
                width=0.1) +
  scale_y_continuous(limits = c(-10,45)) +
  facet_wrap(~location)
```

#### The median counts of each phylum (or species) occurring at each location:
```{r}
ggplot(reef_summary, aes(x=phylum,y=median_count)) +
  geom_point(color="black",
             size=2) +
  geom_errorbar(aes(x=phylum,
                    ymin=median_count - iqr,
                    ymax=median_count + iqr), 
                width=0.1) +
  scale_y_continuous(limits = c(-10,45)) +
  facet_wrap(~location)
```

I will allow people to pick their phylum (or species) of interest and pick their location of interest

### Let's try plotting something else

#### The counts of co-occuring phyla (or species) with a focal phylum (or species)
```{r}
ggplot(reef_phylum, aes(x=fct_rev(phylum))) +
  geom_bar() +
  coord_flip()
```
I will allow people to pick their focal phylum and co-occuring phylum

### Create a table of percent co-occurence of focal genus and neighbor genera (PICK UP TO 3), given total number times focal genus made an appearance
```{r}
#Find number of times focal phylum makes an appearance
reef_focal <- reef_tidy %>%
  filter(binary > "0") %>% 
  mutate(to_match = ifelse(phylum %in% c("cnidaria"), filename, "FALSE")) %>% #create a column that we can subset all rows in a plot based on the presence of focal phylum in the plot at least once
  filter(filename %in% to_match) %>% #if focal phylum is present, keep all observations of that plot ("filename")
  filter(!str_detect(to_match, pattern="FALSE")) %>% #drop rows containing FALSE
  distinct(filename) #get unique plots containing focal phylum

present <- nrow(reef_focal)

#Find number of times neighbor phyla make an appearance
reef_neighbor <- reef_tidy %>%
  filter(binary > "0") %>%
  mutate(to_match = ifelse(phylum %in% c("arthropoda", "chordata", "echinodermata"), filename, "FALSE")) %>%
  filter(filename %in% to_match) %>%
  filter(!str_detect(to_match, pattern="FALSE")) %>% 
  distinct(filename)

absent <- nrow(reef_neighbor)

#Find number of times focal phylum co-occurs with neighbor genus
reef_together <- reef_tidy %>%
  filter(binary > "0") %>% 
  mutate(to_match = ifelse(phylum %in% "cnidaria", filename, "FALSE")) %>%
  filter(filename %in% to_match) %>% 
  mutate(to_match = ifelse(phylum %in% c("arthropoda", "chordata", "echinodermata"), filename, "FALSE")) %>% #do the same thing again for the neighbor phyla
  filter(filename %in% to_match) %>% 
  distinct(filename)

neighborly <- nrow(reef_together)

#Create basic table
reef_table <- as.data.frame(cbind(present, absent, neighborly))%>% 
  mutate(percent_occur = neighborly/present,
         percent_absent = neighborly/absent)

#Generate nice table using `gt`
reef_table %>% 
  gt() %>% 
  fmt_percent(columns=vars(percent_occur, percent_absent), decimal=2) %>% #convert to %
  tab_options(table.width = pct(80)) %>% #make the table width 80% of the page width
  cols_label(present="Focal genus present",
             absent = "Neighbor genera present",
             neighborly="Genera present together",
             percent_occur="Percent co-occurrence",
             percent_absent = "Percent times neighbors occur without focal")
```

Some alternative methods to creating a table
```{r}
df1 <- tribble(
  ~'neighbor', ~'focal', ~'together',  
  nrow(reef_neighbor), nrow(reef_focal), nrow(reef_together)         
) %>%
  mutate(percent_focal = together/focal,
         percent_neighbor = together/neighbor) %>%
  gt() %>% 
  fmt_percent(columns=vars(percent_focal, percent_neighbor), decimal=1) %>% 
  tab_options(table.width = pct(80)) %>% #make the table width 80% of the page width
  cols_label(focal="Focal phylum present",
             neighbor = "Neighbor phyla present",
             together="Phyla present together",
             percent_focal="Percent focal co-occurrs with neighbors",
             percent_neighbor="Percent neighbors co-occurs with focal")

df2 <- as.data.frame(cbind(nrow(reef_focal), nrow(reef_neighbor), nrow(reef_together))) %>% 
  mutate(percent_focal = V3/V1,
         percent_neighbor = V3/V2) %>%
  gt() %>% 
  fmt_percent(columns=vars(percent_focal, percent_neighbor), decimal=1) %>% 
  tab_options(table.width = pct(80)) %>% #make the table width 80% of the page width
  cols_label(V1="Focal phylum present",
             V2 = "Neighbor phyla present",
             V3="Phyla present together",
             percent_focal="Percent focal co-occurrs with neighbors",
             percent_neighbor="Percent neighbors co-occurs with focal")
```

```{r}
reef_summary <- reef_tidy %>%
  filter(binary > "0") %>% #filter out species not present
  st_as_sf(coords=c("longitude", "latitude"), crs=4326) %>% #Make sticky geometries for lat/long
  group_by(location,phylum,orientation) %>% #group by location, then lat/long (and then orientation for vert/horizontal comparisons)
  summarize(`mean abundance` = mean(value), #get the mean count
            median_count = median(value),
            sd_count = sd(value), #get the s.d. count
            iqr = IQR(value), #get the interquartile range for the count
            sample_size = n()) %>% 
  filter(location=="Carp Reef",
         str_detect(orientation,pattern="l")) #add filter to separate vertical from horizontal

library(paletteer)

ggplot(data=reef_summary, aes(x=phylum, y=sample_size)) +
  geom_col() +
  coord_flip() +
  ylab("Abundance") +
  xlab("Phylum") +
  theme_minimal()
```

### Let's make an interactive map! (mean counts of each phylum at each location)
```{r}
reef_sf <- st_as_sf(reef_summary, coords=geometry, crs=4326) #make data sf geometry friendly

reef_map <- tm_basemap("Esri.WorldImagery") +
  tm_shape(reef_sf) + #tells tmap where we're looking (i.e. california)
  #tm_dots(labels="value", col="purple", size=0.5) #tells tmap to plot points on map labelled with id associated
  tm_symbols(id="location", col = "purple", size = "mean_count", scale = 3) +
  tm_facets(by = "phylum")

tmap_mode("view") #view = interactive
reef_map
```

### Let's make an interactive map! (species diversity/richness at each site)
```{r}
reef_index_sf <- st_as_sf(reef_vegan, coords=geometry, crs=4326) #make data sf geometry friendly

reef_map <- tm_basemap("Esri.WorldImagery") +
  tm_shape(reef_index_sf) + #tells tmap where we're looking (i.e. california)
  #tm_dots(labels="value", col="purple", size=0.5) #tells tmap to plot points on map labelled with id associated
  tm_symbols(id="location", col = "diversity", size = "diversity", scale = 3, palette = "inferno", n = 9, contrast = c(1, 0.5)) #view species diversity or richness (BASED ON INPUT)

tmap_mode("view") #view = interactive
reef_map

ggplot(data=reef_index_sf, aes(x=location, y=diversity)) +
  geom_col(aes(fill=diversity)) +
  scale_fill_viridis_c(option = "B", begin = 1, end = 0.5) +
  #scale_fill_brewer(palette="Oranges") +
  #scale_fill_gradient(low="#FFF8C4",high="#B74202") +
  #8E3004 #FFFBD4
  xlab("Location") +
  ylab("diversity") +
  coord_flip() +
  theme_minimal()
```